---
title: "NYC_schools_perceptions"
author: "Bena"
date: "`r Sys.Date()`"
output: html_document
---

# Setting up


```{r}
library(tidyverse)
```

# Import files

The data was downloaded from data.world & NYC education websites using the links below:
  https://data.world/dataquest/nyc-schools-data/workspace/file?filename=combined.csv & 
  https://data.cityofnewyork.us/Education/2011-NYC-School-Survey/mnz3-dyi8 
  
  
```{r}
nyc_hs <- read_csv("combined.csv")
survey_gened <- read_tsv("masterfile11_gened_final.txt")
survey_d75 <- read_tsv("masterfile11_d75_final.txt")
```

# Get overview of data


```{r}
head(nyc_hs)
head(survey_gened)
head(survey_d75)
```


```{r}
dim(nyc_hs)
```

NYC High School data has 479 rows & 30 columns


```{r}
dim(survey_gened)
```

general education survey data has 1646 rows & 1942 columns

```{r}
dim(survey_d75)
```

district75 survey data has 56 rows & 1773 columns

# A closer look at the data

What are the school categories?


```{r}
schooltype_gened <- survey_gened %>%
  pull(schooltype) %>%
  unique

print(schooltype_gened)
```


```{r}
schooltype_d75 <- survey_d75 %>%
  pull(schooltype) %>%
  unique

print(schooltype_d75)
```

School type in d75 data is "District 75 Special Education, which may refer to either high schools or elementary schools.

Since nyc_hs only contains data for NYC high schools, we'll filter the school type for schooltype_gen
We'll also remove some columns & remain with those that are necessary to do the analysis. 
The removed columns show responses to each of the survey questions.


```{r}
survey_gened_select <- survey_gened %>%
  filter(schooltype == "High School") %>%
  select(dbn:aca_tot_11)

View(survey_gened_select)
```


```{r}
survey_d75_select <- survey_d75 %>%
  select(dbn:aca_tot_11)

View(survey_d75_select)
```

Remove the `bn` column since it's contained in `dbn` variable


```{r}
survey_gened_select <- survey_gened_select %>%
  select(-bn)

survey_d75_select <- survey_d75_select %>%
  select(-bn)
```

# Combine the 2 dataframes for surveys


```{r}
survey_combined <- bind_rows(survey_gened_select, survey_d75_select)

View(survey_combined)
```

Alternatively


```{r}
survey_combined <- survey_gened_select %>% 
  bind_rows(survey_d75_select)
```

Since we're interested in relationships of survey_combined data with variables in the nyc_hs dataframe, it makes sense to join the survey data to nyc_hs using left_join().
This will retain only observations in the survey dataframe that correspond to observations in nyc_hs.
Before joining, rename `dbn` so it matches `DBN` in nyc_hs


```{r}
survey_combined <- survey_combined %>%
  rename(DBN = "dbn")
```


```{r}
survey_nychs <- nyc_hs %>%
  left_join(survey_combined, by = "DBN")

View(survey_nychs)
```

# Establish which correlations may be interesting to explore further


```{r}
correlation_matrix <- survey_nychs %>%
  select(avg_sat_score, saf_p_11:aca_tot_11) %>%
  cor(use = "pairwise.complete.obs")

print(correlation_matrix)
```

Convert the matrix to tibble so it's easier to work with


```{r}
correlation_tibble <- correlation_matrix %>%
  as_tibble(rownames = "variable")

print(correlation_tibble)
```

Select variables of interest 
moderate to strong relationships have a correlation coefficient, Pearson's r value <0.25 or >-0.25


```{r}
strong_cors <- correlation_tibble %>%
  select(variable, avg_sat_score) %>%
  filter(avg_sat_score > 0.25 | avg_sat_score < -0.25)

print(strong_cors)
```

# Visualize for closer examination


```{r}
ggplot(data = survey_nychs,
       aes(x = saf_t_11, y = avg_sat_score)) +
       geom_point()
```


```{r}
ggplot(data = survey_nychs,
       aes(x = saf_s_11, y = avg_sat_score)) +
  geom_point()
```


```{r}
ggplot(data = survey_nychs,
       aes(x = aca_s_11, y = avg_sat_score)) +
  geom_point()
```


```{r}
ggplot(data = survey_nychs,
       aes(x = saf_tot_11, y = avg_sat_score)) +
  geom_point()
```

Alternatively, we can iterate using function so we don't have to write 4 different code chunks


```{r}
create_scatter <- function(x, y) {     
  ggplot(data = survey_nychs) + 
    aes_string(x = x, y = y) +
    geom_point(alpha = 0.3) +
    theme(panel.background = element_rect(fill = "white"))
}
x_var <- strong_cors$variable[2:5]
y_var <- "avg_sat_score"

map2(x_var, y_var, create_scatter)

print(x_var)
```

The avg_sat_score has a positive relationship with all 4 variables of interest i.e:
"saf_t_11", "saf_s_11", "aca_s_11", "saf_tot_11"  
In other words, the sat_score increased the more Safety and Respect score (for teachers) increased.
Same case applied to students and all respondent groups.
The sat_score increased the more Academic expectations score (for students) increased.

# Reshape the data so it's more informative

To make it easier to see responses for all metrics by all groups, we'll reshape the data from wide to long data

```{r}
survey_nychs_long <- survey_nychs %>%
  pivot_longer(
    cols = c(saf_p_11:aca_tot_11),       #stacks the perception columns(survey questions) together
    names_to = "survey_question",
    values_to = "score"
  )

View(survey_nychs_long)
```

Create 2 new variables from the survey_question column i.e. metric & respondent


```{r}
survey_nychs_long <- survey_nychs_long %>%
  mutate(
    metric = str_sub(survey_question, 1, 3),       #str_sub() function extracted infor from the survey_question variable
    respondent = str_sub(survey_question, 4, 6)    # start & end characters could have been 5, 5 but there are 2 t's for teacher & total
  )
```

Replace the values in metric & respondent columns with more meaningful names

```{r}
survey_nychs_long <- survey_nychs_long %>%
  mutate(
    metric = case_when(
      metric == "saf" ~ "safety & respect",
      metric == "com" ~ "communication",
      metric == "eng" ~ "engagement",
      metric == "aca" ~ "academic expectations"
    )
  )
```


```{r}
survey_nychs_long <- survey_nychs_long %>%
  mutate(
    respondent = case_when(
      respondent == "_p_" ~ "parent",
      respondent == "_t_" ~ "teacher",
      respondent == "_s_" ~ "student",
      respondent == "_to" ~ "total"
    )
  )
```

Alternatively, if_else() can be used in place of case_when().
Restore data in original long format then use if_else()


```{r}
survey_nychs_long <- survey_nychs %>%
  pivot_longer(
    cols = c(saf_p_11:aca_tot_11),       #stacks the perception columns(survey questions) together
    names_to = "survey_question",
    values_to = "score"
  )
```


```{r}
survey_nychs_long <- survey_nychs_long %>%
  mutate(
    metric = str_sub(survey_question, 1, 3),       #str_sub() function extracted infor from the survey_question variable
    respondent = str_sub(survey_question, 4, 6)    # start & end characters could have been 5, 5 but there are 2 t's for teacher & total
  )
```


```{r}
survey_nychs_long <- survey_nychs_long %>%
  mutate(metric = ifelse(metric  == "saf", "safety & respect", 
                             ifelse(metric == "com", "communication",
                                    ifelse(metric == "eng", "engagement", 
                                           ifelse(metric == "aca", "academic expectations", "NA")))))
```


```{r}
survey_nychs_long <- survey_nychs_long %>%
  mutate(respondent = ifelse(respondent  == "_p_", "parent", 
                                ifelse(respondent == "_t_", "teacher",
                                       ifelse(respondent == "_s_", "student", 
                                              ifelse(respondent == "_to", "total", "NA")))))
```

Now it's manageable to visualize the perceptions


```{r}
ggplot(data = survey_nychs_long,                    #visualizes all responses
       aes(x = metric, y = score, color = respondent)) +
       geom_boxplot()
```


```{r}
survey_nychs_long %>%
  filter(respondent != "total") %>%                 
  ggplot(aes(x = metric, y = score, fill = respondent)) +
  geom_boxplot()
```


From the visualization created with code above, it was noted that:

- Scores for teachers were more spread out than those for students and parents
- The larger & smaller scores are equally spread out for all metrics across all groups except safety & respect(among parents) and academic expectations (among students)
- For academic expectations(among students), those that are smaller than medium value are close together.
- For safety & respect(among parents), those that are higher than medium value are closer together.

# Other codes


```{r}
HS <- survey_combined %>%   #checking unique entries for highschool column
  pull(highschool) %>%
  unique

print(HS)
```


```{r}
surveyed <- survey_combined %>%   #checking unique entries for studentssurveyed variable
  pull(studentssurveyed) %>%
  unique

print(surveyed)
```


Investigate those which were not surveyed


```{r}
not_surveyed <- survey_combined %>%
  filter(studentssurveyed == "No")

View(not_surveyed)  #Just 1 obs.
```



